---
title: "Report Exercise 3"
author: "Michele Iannuzzo"
date: "`r Sys.Date()`"
output: html_document
---

Load the necessary libraries.
```{r}
library(tidyverse)
library(dplyr)
library(knitr)
```


In a first step, the data has been cleaned manually. The decomposition data file is the cleaned file.

Now load the data and get rid of unnecessary columns and rows.
```{r}
decomposition <- read_delim("../data/decomposition.csv", delim = ";") #use read_delim so I can specify the delimiter

decomposition <- subset(decomposition, select=c(1:15)) #remove unnecessary columns from dataframe
decomposition <- decomposition[c(1:234),] #remove unnecessary rows
decomposition #to see dataframe in console
```
The next step is to aggregate data per experiment.
I herefore use the aggregate() function on the "decomposition" data frame. I list() the data frame according to the "Experiment" column and use the function FUN = mean so R will take the mean for each aggregate.
Finally, I present the results as a table with the kable() function.
```{r, warning=FALSE}
#aggregate data per experiment
aggregate_mean <- aggregate(decomposition,
                         by = list(decomposition$Experiment),
                         FUN = mean)
knitr::kable(aggregate_mean)
```
Then I calculate the Log-respose ratio within each experiment.
Therefore, I use the previously saved aggregates ("aggregate_mean"), group them with group_by(). "Group.1" are the different Experiments that were aggregated in the step before. As a last step, I use summarise() with the given formula to calcuolate the logarithmic response ratio. 
Finally, I present the results as a table with the kable() function.
```{r}
#Log-response ratio
decomposition_log <- aggregate_mean |>
  group_by(Group.1) |>
  summarise(RR = log(`mean increased CO2 [gCm-2]`/`mean ambient CO2 [gCm-2]`))

knitr::kable(decomposition_log)
```

Now, I am going to aggregate data across all experiments for different years.
As a first step, I have to mutate() the Time stamp, so I can aggregate the data according to the years. In a second step, I use the new data frame "agg_decomposition" with now in tidyverse readable years. I mutate() them again to assign all data points to one of the three time periods (<3 years, 3-6 years, >=6 years). In a last step, I aggregate the data by those three time periods ("phase"), again, with the function "mean", so R will take the mean over the phases.
Finally, I present the results as a table with the kable() function.
```{r, warning=FALSE}
#Aggregate date per year 
agg_decomposition <- decomposition |>
  mutate(`Time (years)` = as.numeric(`Time (years)`))
str(decomposition$`Time (years)`)

agg_decomposition <- agg_decomposition |>
  mutate(phase = if_else(agg_decomposition$`Time (years)` < 3,"<3 years", 
                         if_else(agg_decomposition$`Time (years)` >=6, ">=6 years", "3-6 years")))

agg_decomposition <- aggregate(agg_decomposition,
                         by = list(agg_decomposition$phase),
                         FUN = mean)

knitr::kable(agg_decomposition)
```

And for the last step I will calculate the log-response ratio for each phase.
As in the previous log-response calculation, I group by the three time periods and summarise() with the given formula.
And finally, I present the results as a table with the kable() function.
```{r}
#Log-response ratio for each phase
RR_decomposition <- agg_decomposition |>
  group_by(Group.1) |>
  summarise(RR = log(`mean increased CO2 [gCm-2]`/`mean ambient CO2 [gCm-2]`))

knitr::kable(RR_decomposition)
```


